---
title: "Baby Names Graph"
author: "Chris Moroney"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r load-packages, message=FALSE}
library(mdsr)
library(tidyverse)
library(babynames)
library(Hmisc)
```


### Original code from MDSR

```{r}
BabynamesDist <- make_babynames_dist()
head(BabynamesDist, 2) 

BabynamesDist %>% filter(name == "Benjamin") 

Joseph <- BabynamesDist %>%  
  filter(name == "Joseph" & sex == "M")  
name_plot <- ggplot(data = Joseph, aes(x = year)) 

name_plot <- name_plot +  
  geom_bar(stat = "identity", aes(y = count_thousands * alive_prob),  
           fill = "#b2d7e9", colour = "white") 

name_plot <- name_plot + geom_line(aes(y = count_thousands), size = 2) 

name_plot <- name_plot +  
  ylab("Number of People (thousands)") + xlab(NULL) 

summary(name_plot) 

wtd.quantile <- Hmisc::wtd.quantile  
median_yob <-  
  with(Joseph, wtd.quantile(year, est_alive_today, probs = 0.5)) 
median_yob 

name_plot <- name_plot +
  geom_bar(stat = "identity", colour = "white", fill = "#008fd5",
           aes(y = ifelse(year == median_yob, est_alive_today/ 1000, 0))) 

name_plot +  
  ggtitle("Age Distribution of American Boys Named Joseph") +  
  geom_text(x = 1935, y = 40, label = "Number of Josephs\nborn each year") + 
  
  geom_text(x = 1915, y = 13, label = 
              "Number of Josephs\nborn each year\nestimated to be alive\non 1/1/2014",  
            colour = "#b2d7e9") +  
  geom_text(x = 2003, y = 40,  
            label = "The median\nliving Joseph\nis 49 years old",  colour = "darkgray") + 
  geom_curve(x = 1995, xend = 1974, y = 40, yend = 24,  
             arrow = arrow(length = unit(0.3,"cm")), curvature = 0.5) + ylim(0, 42) 
```

### Reproduced Code with "Chris" instead of Joseph and Code from book

Here is all of the code and modified for my name, which is Chris

```{r code-chunk-label}
BabynamesDist <- make_babynames_dist()
head(BabynamesDist, 2) 

BabynamesDist %>% filter(name == "Chris") 

Chris <- BabynamesDist %>%  
  filter(name == "Chris" & sex == "M")  
name_plot <- ggplot(data = Chris, aes(x = year)) 

name_plot <- name_plot +  
  geom_bar(stat = "identity", aes(y = count_thousands * alive_prob),  
           fill = "#b2d7e9", colour = "white") 

name_plot <- name_plot + geom_line(aes(y = count_thousands), size = 2) 

name_plot <- name_plot +  
  ylab("Number of People (thousands)") + xlab(NULL) 

summary(name_plot) 

wtd.quantile <- Hmisc::wtd.quantile  
median_yob <-  
  with(Chris, wtd.quantile(year, est_alive_today, probs = 0.5)) 
median_yob 

name_plot <- name_plot +
  geom_bar(stat = "identity", colour = "white", fill = "#008fd5",
           aes(y = ifelse(year == median_yob, est_alive_today/ 1000, 0))) 

name_plot +  
  ggtitle("Age Distribution of American Boys Named Chris") +  
  geom_text(x = 1935, y = 8, label = "Number of Chris's\nborn each year") + 
  
  geom_text(x = 1920, y = 2.5, label = 
              "Number of Chris's\nborn each year\nestimated to be alive\non 1/1/2014",  
            colour = "#b2d7e9") +  
  geom_text(x = 1993, y = 7.5,  
            label = "The median\nliving Chris\nis 49 years old",  colour = "darkgray") + 
  geom_curve(x = 1985, xend = 1965, y = 7.5, yend = 5.75,  
             arrow = arrow(length = unit(0.3,"cm")), curvature = 0.5) + ylim(0, 10) 
```


### List of Graph changes

- Replaced all instances of “Joseph” with “Chris” or “Chris’s”
- Changed the Label location of "number of Chris's\nborn each year" to x=1935, y=8
- Changed the Label location of "Number of Chris's\nborn each year\nestimated to be alive\non 1/1/2014" to x = 1920, y = 2.5
- Changed the Label location of The median\nliving Chris\nis 49 years old" to x=1993, y=7.5
- Changed the arrow coordinates to be x=1985, xend=1965, y=7.5, yend=5.75, 
- Changed the scale of the y-axis to limit from 40 thousand to 10 thousand, which became ylim(0, 10)